\documentclass[11pt,a4paper]{article}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage{geometry}
\usepackage{enumitem}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{tikz}
\usetikzlibrary{automata,positioning,arrows.meta}
\geometry{margin=2.5cm}

\tikzset{
  kairosstate/.style={
    state,
    minimum width=20mm,
    minimum height=9mm,
    inner sep=2pt,
    font=\small
  },
  kairoslabel/.style={
    font=\small,
    align=center,
    inner sep=1pt
  }
}

\newtheorem{definition}{Definition}[section]
\newtheorem{proposition}[definition]{Proposition}
\newtheorem{theorem}[definition]{Theorem}
\newtheorem{lemma}[definition]{Lemma}
\newtheorem{remark}[definition]{Remark}

\newcommand{\Nat}{\mathbb{N}}
\newcommand{\Stream}[1]{#1^\omega}
\newcommand{\True}{\mathsf{true}}
\newcommand{\False}{\mathsf{false}}
\newcommand{\StepRel}{\mathsf{Step}}
\newcommand{\Cfg}{\mathsf{Cfg}}
\newcommand{\Obs}{\mathsf{Obs}}
\newcommand{\TickCtx}{\mathsf{TickCtx}}
\newcommand{\UserInv}{\mathsf{Inv}}
\newcommand{\Support}{\mathsf{Support}}
\newcommand{\Match}{\mathsf{Match}}
\newcommand{\BadLocal}{\mathsf{BadLocal}}
\newcommand{\GeneratedBy}{\mathsf{GeneratedBy}}
\newcommand{\Generated}{\mathsf{Generated}}
\newcommand{\GeneratedTriple}{\mathsf{GeneratedTriple}}
\newcommand{\AvoidA}{\mathsf{AvoidA}}
\newcommand{\AvoidG}{\mathsf{AvoidG}}
\newcommand{\run}{\mathsf{run}}
\newcommand{\rps}{\mathsf{rps}}
\newcommand{\ctx}{\mathsf{ctx}}
\newcommand{\head}{\mathsf{hd}}
\newcommand{\tail}{\mathsf{tl}}
\newcommand{\ResetOk}{\mathsf{ResetOk}}
\newcommand{\GuardReset}{\mathsf{GRst}}
\newcommand{\GuardAfterReset}{\mathsf{GAfterRst}}
\newcommand{\GuardDelay}{\mathsf{GDelay}}

\title{Conditional Safety, Explicit Products, and Relational Hoare Triples for Kairos}
\author{Kairos Project}
\date{\today}

\begin{document}
\maketitle

\begin{abstract}
Kairos targets conditional safety properties of reactive programs over infinite
streams: every admissible input stream should produce an execution satisfying a
safety specification. The challenge is that the property is global in time,
whereas the proof objects used in practice must remain local and compatible
with deductive reasoning over programs with unbounded state. We present a
semantic decomposition of this problem based on an explicit product between the
program and two finite safety monitors representing assumptions and guarantees.
From this product, we derive dangerous local steps, generated semantic clauses,
and relational Hoare triples over reactive ticks. The main theorem states that,
under explicit well-formedness hypotheses on the program model, validity of all
generated triples suffices for conditional safety. The paper gives a rigorous
account of the reactive semantics, the clause-to-triple discipline, and the
proof structure that reduces a global safety violation to a contradiction with
a valid local triple. Backend grouping, encoding, and solver interaction are
treated separately as refinements of this mathematical core.
\end{abstract}

\section{Introduction}
Synchronous verification has long relied on stream semantics, observers,
contracts, and theorem proving. In Lustre-like settings, one reasons about
reactive components that evolve in lockstep on infinite streams, either through
observers and synchronous contracts or through proof-oriented semantic
encodings~\cite{lustre-pvs,agree-techreport,agree-observers,kind2-realizability,cocospec}.
In parallel, another line of work compiles temporal specifications into local
program proof objects that can be discharged by deductive backends, as in
Aorai or CaFE for C programs~\cite{aorai-manual,stouls-groslambert-2009,cafe-page}.
Kairos lies at the intersection of these two traditions: it is synchronous and
reactive in the sense of stream-based semantics, but it also seeks a reduction
from global temporal safety to local Hoare-style reasoning over ticks. This
combination immediately raises the methodological question that drives the
paper: what is the right semantic interface between a stream-level safety
property and transition-level deductive reasoning?

The setting considered here is conditional safety over reactive executions. A
program runs under an environment hypothesis on input streams, and its
correctness is expressed by a safety guarantee on the induced input/output
trace. The user may also provide node invariants that capture additional facts
about executions. The resulting statement is neither a plain state invariant nor
a mere operational contract: it is a theorem of the form ``admissible inputs
imply safe executions,'' where both admissibility and safety are understood as
properties of infinite traces.

The main obstacle is not merely temporal expressiveness. The programs of
interest manipulate non-bounded data domains: memories, inputs, and outputs are
not assumed finite-state. This rules out a direct explicit-state account of the
whole verification problem. The core challenge is therefore to preserve the
meaning of a global safety property while reducing its proof to local objects
that remain compatible with deductive reasoning over an unbounded transition
system.

This question is easy to underestimate. If one reasons only with finite
temporal monitors, one forgets the unbounded program state. If one reasons only
with Hoare-style local contracts, one loses the global temporal meaning of the
specification. A paper at the level expected by POPL or PLDI must therefore
make explicit not only the final theorem, but also the exact semantic path from
global temporal safety to local deductive proof objects. This is the role of
the present development.

The core proof objects of the paper are relational Hoare triples over reactive
ticks. Each triple speaks about one program transition, a precondition on the
current local tick context, and a postcondition on the successor context. The
problem is then to generate the right preconditions and postconditions so that
validity of these local triples really implies the original global safety
statement.

To solve this problem, we exploit the fact that standard safety specification
formalisms can be compiled to finite safety automata; this includes, in
particular, the safety fragment of linear temporal logic used by the current
Kairos toolchain~\cite{kupferman-vardi-safety,gastin-oddoux}. The paper
therefore presents the mathematics directly in terms of assumption and
guarantee safety automata. This allows us to define an explicit semantic
product \(program \times A \times G\), where \(A\) is the assumption automaton
and \(G\) is the guarantee automaton. The product synchronizes, at each tick, a
program transition, the reading of the current input by \(A\), and the reading
of the current observation by \(G\). It provides the right semantic granularity
for identifying dangerous local steps, extracting semantic clauses, and
assembling the generated relational Hoare triples.

\paragraph{Contributions.}
The paper makes four contributions.
\begin{enumerate}[leftmargin=1.4em]
\item It defines a mathematical model of reactive programs over streams,
  together with a separate specification layer made of an assumption automaton,
  a guarantee automaton, and node invariants over local trace contexts.
\item It introduces the explicit product \(program \times A \times G\) as the
  central semantic object from which dangerous steps and generated clauses are
  defined.
\item It formalizes a clause-to-triple discipline: generated semantic clauses
  are not the final proof objects; they are assembled into relational Hoare
  triples whose validity suffices for the conditional safety theorem.
\item It reconstructs a readable proof architecture, aligned with the Rocq
  development, that derives the global theorem from product progression,
  helper propagation, clause coverage, and triple coverage.
\end{enumerate}

\paragraph{What is non-standard here.}
The novelty is not the mere use of automata, nor the mere use of deductive
verification. The non-standard point is the way these two ingredients are tied
together. The product \(program \times A \times G\) is not introduced as an
implementation device, nor as a model-checking state space to be exhaustively
explored. It is the semantic object that justifies clause extraction, triple
generation, and the reduction from a global bad execution to a local
contradiction.

\paragraph{Scope of the result.}
The core theorem developed in this paper stops at semantic validity of
generated relational Hoare triples. It does not prove the correctness of the
backend grouping strategy, of the Why3 encoding, or of external prover
integration. These are refinement layers that matter for the tool, but they are
intentionally kept outside the core mathematical statement.

The remainder of the paper is organized as follows. Section~2 defines reactive
programs, their stream semantics, and the local trace contexts on which
specifications are interpreted. Section~3 introduces deterministic safety
automata and the interpretation of local predicates, including predicates with
past. Section~4 presents the explicit product and its semantic role. Section~5
defines dangerous steps, local contexts, and the matching relation. Section~6
describes clause generation and triple assembly in a pseudo-algorithmic yet
formal way. Section~7 reconstructs the proof chain from a global bad execution
to the contradiction yielded by a valid generated triple. Section~8 discusses
the connection between the mathematical presentation and the current
implementation architecture.

\section{Reactive Program and Stream Semantics}

\subsection{Reactive Program}

A Kairos program is viewed as a stateful reactive system. The program state
contains a control component and a memory component. Program transitions are not
written as imperative assignments, but as logical relations between the current
state, the current input, the current output, and the next state.

\begin{definition}[Reactive Program]
A reactive program \(P\) is given by:
\[
P = (S,M,I,O,T,s_0,m_0,\StepRel),
\]
where:
\begin{itemize}[leftmargin=1.4em]
\item \(S\) is the set of control states;
\item \(M\) is the set of memory states;
\item \(I\) is the set of inputs;
\item \(O\) is the set of outputs;
\item \(T\) is a set of transition identifiers;
\item \(s_0 \in S\) and \(m_0 \in M\) are the initial conditions;
\item \(\StepRel(t,s,m,i,s',m',o)\) is a logical predicate, read as:
  transition \(t\) allows the passage from \((s,m)\) to \((s',m')\) under
  input \(i\), producing output \(o\).
\end{itemize}
\end{definition}

\begin{definition}[Program Specification]
The specification associated with \(P\) is a triple
\[
\Phi = (A,G,\UserInv),
\]
where:
\begin{itemize}[leftmargin=1.4em]
\item \(A\) is a safety automaton over input streams;
\item \(G\) is a safety automaton over reactive input/output traces;
\item \(\UserInv : S \to (\TickCtx \to \mathsf{Prop})\) is an aggregated node
  invariant: for each control state \(s\), \(\UserInv(s)\) is a property over
  tick contexts compatible with the execution.
\end{itemize}
\end{definition}

\begin{remark}
The key point is the distinction between:
\begin{itemize}[leftmargin=1.4em]
\item the \emph{program} \(P\), which contains only its syntax and step
  semantics;
\item the \emph{specification} \(\Phi=(A,G,\UserInv)\), which contains the
  external hypotheses, the guarantee, and node formulas over the execution
  trace.
\end{itemize}
Kairos clauses therefore do not arise from an ``enriched program'' object,
but from the coupling of \(P\) and \(\Phi\).
\end{remark}

\subsection{Determinism and Step Relation}

To avoid unnecessary overhead, we work in the deterministic setting targeted by
Kairos: at each tick, the program selects a unique transition, and the automata
also select a unique outgoing edge.

\begin{definition}[Deterministic Program Step]
Assume a functional step relation such that, for every
\((s,m,i)\), there exists a unique triple \((t,(s',m'),o)\) satisfying
\[
((s,m),i)\xrightarrow{\,t\,}_P((s',m'),o).
\]
This relation is compatible with \(\StepRel\), namely:
\[
((s,m),i)\xrightarrow{\,t\,}_P((s',m'),o)
\Longrightarrow
\StepRel(t,s,m,i,s',m',o).
\]
\end{definition}

\begin{remark}[Totality Assumption on Program Ticks]
The abstract model used in this paper assumes that every instant is total: from
every current configuration and input valuation, the program computes a next
control state, a next memory, and an output valuation. This is stronger than
what a concrete implementation language may guarantee by construction. In
particular, a concrete program may block because of an invalid operation, or
fail to emit a value on some output signal. Such situations are intentionally
abstracted away from the mathematical core and must therefore be ruled out by
separate well-formedness conditions when one applies the model to actual Kairos
programs.
\end{remark}

\begin{remark}[Exhaustiveness of Transition Filtering]
Another practical source of non-totality is the absence of any active
transition for the current control state, memory, and input valuation. In the
abstract model, this case is excluded: the filtering of transitions by current
state and guard satisfaction is assumed to be exhaustive on the executions of
interest. In practice, this should be guaranteed structurally, for instance by
requiring exhaustive guarded cases or by inserting an explicit default branch
at the source-language level. Introducing a semantic sink state inside the
abstract program model would require specifying its observable output, which is
not neutral; for this reason, the present paper treats exhaustiveness as a
well-formedness requirement on concrete programs rather than as part of the
core semantics.
\end{remark}

\subsection{Stream Semantics}

A local transition is not enough: one must define the execution of the program
over an infinite input stream. This is where stream semantics comes in.

\begin{definition}[Stream]
For every set \(X\), we write
\[
\Stream{X} := \Nat \to X
\]
for the set of infinite streams over \(X\).
\end{definition}

\begin{definition}[Configuration and Observation]
We write
\[
\Cfg := S \times M,
\qquad
\Obs := I \times O.
\]
\end{definition}

\begin{definition}[Admissible Initial Memories]
We assume a predicate
\[
\mathsf{InitMem} \subseteq M
\]
of admissible initial memories. The program starts in the distinguished control
state \(s_0\), but the initial memory is not required to be a single known
value: any \(m_0 \in \mathsf{InitMem}\) is allowed.
\end{definition}

\begin{definition}[Local Tick Context]
We define
\[
\TickCtx := \Nat \times \Cfg \times I \times O \times \Cfg.
\]
An element
\[
(k,(s,m),i,o,(s',m')) \in \TickCtx
\]
is called a \emph{local tick context}. It represents the smallest execution
window around tick \(k\): current configuration, current input, current output,
and next configuration.
\end{definition}

\begin{definition}[Trace-Derived Local Context]
Fix an admissible initial memory \(m_0 \in \mathsf{InitMem}\) and an input stream
\(u \in \Stream{I}\), and let \((s_{m_0,u},m_{m_0,u},o_{m_0,u})\) be the
execution generated by \(P\) from \(m_0\) on \(u\). We write
\[
\ctx_{m_0,u}(k) :=
\bigl(k,\,(s_{m_0,u}(k),m_{m_0,u}(k)),\,u(k),\,o_{m_0,u}(k),\,
      (s_{m_0,u}(k+1),m_{m_0,u}(k+1))\bigr)
\in \TickCtx,
\]
that is, the context obtained at tick \(k\) from the current configuration, the
current input, the current output, and the next configuration of the execution.
\end{definition}

\paragraph{Notation conventions.}
Fix an admissible initial memory \(m_0 \in \mathsf{InitMem}\) and an input stream
\(u \in \Stream{I}\). The program \(P\) then generates an execution on
\((m_0,u)\), namely the sequences of control states, memories, and outputs
produced tick after tick by the transition relation of \(P\). We
systematically write
\[
s_{m_0,u} : \Nat \to S,\qquad m_{m_0,u} : \Nat \to M,\qquad o_{m_0,u} : \Nat \to O
\]
for these three components of the execution of \(P\) from \(m_0\) on \(u\). To
keep formulas readable, we omit the index \(m_0\) when it is fixed by context
and simply write \(s_u\), \(m_u\), \(o_u\), \(\ctx(u,k)\), and \(\run(u)\).
When we need to make the initial memory explicit, we write
\(\ctx_{m_0,u}(k)\) and \(\run_{m_0}(u)\).
Thus:
\begin{itemize}[leftmargin=1.4em]
\item \(s_0\) denotes the distinguished initial control state of the program;
\item \(m_0 \in \mathsf{InitMem}\) denotes an admissible initial memory;
\item \(s_u(k)\) and \(m_u(k)\) denote the control state and memory reached at
  tick \(k\) under input \(u\);
\item \(o_u(k)\) denotes the output produced at tick \(k\).
\end{itemize}

For an input \(u \in \Stream{I}\), program execution must produce:
\begin{itemize}[leftmargin=1.4em]
\item a sequence of configurations \((s_u(k),m_u(k))\);
\item a sequence of outputs \(o_u(k)\);
\item a reactive trace \(\run(u)(k) = (u(k),o_u(k))\).
\end{itemize}

\paragraph{Primary definition.}
In this document, stream semantics is defined pointwise first.

\begin{definition}[Stream Semantics]
For every admissible initial memory \(m_0 \in \mathsf{InitMem}\) and every input
stream \(u \in \Stream{I}\), we define the sequences \((s_u,m_u,o_u)\) by:
\[
(s_u(0),m_u(0)) = (s_0,m_0),
\]
and, for every \(k \in \Nat\), by the unique deterministic step
\[
((s_u(k),m_u(k)),u(k)) \xrightarrow{\,t_k\,}_P ((s_u(k+1),m_u(k+1)),o_u(k)).
\]
The reactive trace is then defined by
\[
\run(u)(k) := (u(k),o_u(k)).
\]
\end{definition}

\begin{remark}[Coinductive Reading]
This pointwise definition corresponds to the standard coinductive reading of a
guarded stream equation: the head of the stream is given by the first program
step, and the tail is obtained by reapplying the same rule to the input suffix
and the next configuration. We use this reading as intuition, but the primary
presentation in the paper is intentionally the operational definition above.
\end{remark}

\section{Safety Automata and Predicate Interpretation}

\subsection{Safety Automata}

The specification \(\Phi=(A,G,\UserInv)\) contains three components: an
assumption automaton \(A\), a guarantee automaton \(G\), and a node invariant
\(\UserInv\). We begin by fixing the general semantics of \(A\) and \(G\).

A safety automaton reads an infinite word and rejects as soon as it reaches a
\(\mathit{bad}\) state. In Kairos, edge labels are also written as logical
relations, not as commands.

\begin{definition}[Deterministic Safety Automaton]
A deterministic safety automaton over an observation domain \(X\) is a
tuple
\[
\mathcal{A} = (Q,q_0,bad,E,\mathsf{src},\mathsf{dst},\mathsf{Lab},\mathsf{select}),
\]
where:
\begin{itemize}[leftmargin=1.4em]
\item \(Q\) is the set of states;
\item \(q_0 \in Q\) is the initial state;
\item \(bad \in Q\) is the violation state;
\item \(E\) is a set of edges;
\item \(\mathsf{src},\mathsf{dst} : E \to Q\) give the source and target;
\item \(\mathsf{Lab}(e) : X \to \mathsf{Prop}\) is the logical edge predicate;
\item \(\mathsf{select}(q,x)\) selects the unique outgoing edge from \(q\)
  whose label holds on \(x\).
\end{itemize}
\end{definition}

\begin{definition}[Run and Acceptance Condition]
If \(w \in \Stream{X}\), the execution of \(\mathcal{A}\) on \(w\) is defined by
\[
q_{\mathcal{A},w}(0)=q_0,
\qquad
q_{\mathcal{A},w}(k+1)
=
\mathsf{dst}\bigl(\mathsf{select}(q_{\mathcal{A},w}(k),w(k))\bigr).
\]
The word \(w\) is accepted as safe if
\[
\forall k \in \Nat,\qquad q_{\mathcal{A},w}(k) \neq bad.
\]
\end{definition}

\subsection{Interpretation of Local Predicates}

The formulas used in \(A\), \(G\), and \(\UserInv\) are interpreted over
\(\ctx_u(k)\) contexts, not directly over the raw memory.

\begin{definition}[Variables and Current Evaluation]
Assume a set \(\mathsf{Var}\) of observable program variables, together with a
current-value function
\[
\mathsf{val}_{\mathit{now}} : \TickCtx \times \mathsf{Var} \to \mathsf{Val}
\]
that reads the current value of each variable in the tick context. A free
variable \(x\) in a formula therefore means the current value of \(x\) in the
context, provided that \(x\) belongs to the domain of allowed observable
variables.
\end{definition}

\begin{remark}[Abstract History Interface]
The abstract model does not require a distinguished syntax of past operators.
It only assumes that labels and invariants are predicates over local trace
contexts and may therefore depend on suitable functions of the past. Concrete
constructs such as \(prev\) or \(pre_k\) should be read as examples of how one
may instantiate this abstract interface, not as primitive constituents of the
core model.
\end{remark}

\begin{definition}[Example Instantiation of History]
If \(x \in \mathsf{Var}\), one possible instantiation interprets \(prev(x)\) at
tick \(k>0\) by
\[
\mathsf{eval}_{\ctx_u(k)}(prev(x)) :=
\mathsf{val}_{\mathit{now}}(\ctx_u(k-1),x).
\]
More generally, \(pre_k(x)\) is interpreted by reading at tick \(k-d\), whenever
it is defined. Variables not placed under a past operator remain interpreted by
the current valuation
\(\mathsf{val}_{\mathit{now}}(\ctx_u(k),x)\).
\end{definition}

\begin{remark}
The user-level intuition is that one may write formulas over current variables
and, in a concrete language instance, over derived history operators such as
\(prev\). The abstract semantics only fixes how such formulas are interpreted
through trace contexts. The finite encoding through auxiliary variables only
appears later, on the implementation side.
\end{remark}

\subsection{Node Invariant}

\begin{definition}[Node Invariant]
A node invariant is a function
\[
\UserInv : S \to (\TickCtx \to \mathsf{Prop}).
\]
Thus, for each control state \(s\), \(\UserInv(s)\) is a single property over
tick contexts. If the user provides several elementary invariants for the same
state, they are aggregated by conjunction into this single property.
\end{definition}

\begin{remark}
This matches the current Rocq presentation, where the node invariant is indexed
by the state and then evaluated over a trace context. Semantically, this is
cleaner than taking a primitive collection of distinct properties.
\end{remark}

\subsection{Example: \texttt{resettable\_delay}}

The running example is a resettable one-tick delay. It has two control states
\[
S = \{Init,Run\},
\]
with memory \(M = \mathbb{Z}\), input domain
\[
I = \mathbb{B} \times \mathbb{Z},
\]
where the two input components are written \(reset\) and \(x\), output
\(y \in O = \mathbb{Z}\), and the following four relational transitions:
\begin{align*}
\StepRel(t_{init}^{rst},Init,m,(true,x),Run,m',y)
&:\Longleftrightarrow
y = 0 \land m' = 0,\\
\StepRel(t_{init}^{keep},Init,m,(false,x),Run,m',y)
&:\Longleftrightarrow
y = 0 \land m' = x.
\end{align*}
and, in the steady state,
\begin{align*}
\StepRel(t_{run}^{rst},Run,m,(true,x),Run,m',y)
&:\Longleftrightarrow
y = 0 \land m' = 0,\\
\StepRel(t_{run}^{keep},Run,m,(false,x),Run,m',y)
&:\Longleftrightarrow
y = m \land m' = x.
\end{align*}

Intuitively, the node behaves as follows. If \(reset\) is true, it outputs
\(0\) and clears its memory. Otherwise, it behaves like an ordinary one-tick
delay: it outputs the current memory and stores the current sample \(x\). We
use this example as the running case study because it remains compact while
exhibiting a genuine interaction between input assumptions, guarantee clauses,
and user invariants.

A natural user annotation is the aggregated node invariant
\[
\UserInv(Run)(\ctx_u(k)) :\Longleftrightarrow
\bigl(prev(reset) \Rightarrow m = 0\bigr)
\;\land\;
\bigl(\neg prev(reset) \Rightarrow m = prev(x)\bigr),
\]
where \(m\) denotes the unique memory cell. Since \(M=\mathbb{Z}\) in this
example, memory is reduced to a single integer register. The invariant says
that, in state \(Run\), the memory records either the reset value \(0\) or the
previous sample, depending on whether the preceding tick asserted \(reset\).
Unfolded on the execution:
\[
\psi_{rdelay}(k) :\Longleftrightarrow
s_u(k)=Run \Rightarrow
\bigl(prev(reset) \Rightarrow m_u(k)=0\bigr)
\land
\bigl(\neg prev(reset) \Rightarrow m_u(k)=prev(x)\bigr),
\qquad (k \ge 1).
\]

\begin{figure}[h]
\centering
\begin{tikzpicture}[>=Stealth,semithick]
  \node[kairosstate,initial] (init) {$Init$};
  \node[kairosstate,right=58mm of init] (run) {$Run$};
  \path[->]
    (init) edge[bend left=15] node[kairoslabel,midway,above]
      {$reset$\\$y = 0$\\$m' = 0$} (run)
    (init) edge[bend right=15] node[kairoslabel,midway,below]
      {$\neg reset$\\$y = 0$\\$m' = x$} (run)
    (run) edge[loop above] node[kairoslabel]
      {$reset$\\$y = 0$\\$m' = 0$}
      (run)
    (run) edge[loop right] node[kairoslabel]
      {$\neg reset$\\$y = m$\\$m' = x$}
      (run);
\end{tikzpicture}
\caption{Program automaton for \texttt{resettable\_delay}.}
\end{figure}

This program is simple enough to be understood operationally, but rich enough
to require both assumption-side and invariant-side reasoning in the local proof
objects.

\section{Specification, Observations, and Guarantees}

\subsection{From Alphabet Words to Semantic Observations}

Automaton labels do not necessarily range over a finite syntactic alphabet.
They may instead be read as predicates over semantic observations.

\begin{definition}[Observation Domains]
In Kairos, one uses in practice:
\begin{itemize}[leftmargin=1.4em]
\item \(X_A = I\) for the assumption automaton \(A\);
\item \(X_G = I \times O\) for the guarantee automaton \(G\);
\item more generally, a tick context derived from the trace if labels mention
  history.
\end{itemize}
\end{definition}

\begin{remark}
The presence of past atoms such as \(prev(x)\) does not change the nature of
the automaton. It only means that the label is evaluated on the trace through a
local context \(\ctx_u(k)\) rich enough to interpret the past. In the
implementation, this reading is then compiled through finite auxiliary
variables; these variables are not part of the source semantics.
\end{remark}

\subsection{Running Example: Assumption and Guarantee Automata}

For the running \texttt{resettable\_delay} example, we consider the following
conditional safety specification:
\[
A := G(\ResetOk),
\]
and
\[
G :=
G(\GuardReset)
\;\land\;
X\,G(\GuardAfterReset)
\;\land\;
X\,G(\GuardDelay).
\]
where the three local predicates are
\[
\ResetOk := (\neg reset \vee x = 0),
\]
\[
\GuardReset := (reset \Rightarrow y = x),
\]
\[
\GuardAfterReset := (\neg reset \land prev(reset) \Rightarrow y = 0),
\]
\[
\GuardDelay := (\neg reset \land \neg prev(reset) \Rightarrow y = prev(x)).
\]
The assumption says that a reset command is only issued together with the value
\(x=0\). The guarantee has three branches:
\begin{itemize}[leftmargin=1.4em]
\item on a reset tick, the output must agree with the reset value;
\item on the first non-reset tick after a reset, the output must still be
  \(0\);
\item on ordinary non-reset operation, the output must equal the previous
  sample.
\end{itemize}

The assumption automaton has only two states:
\[
Asm_{ok},\qquad Asm_{bad}.
\]

\begin{figure}[h]
\centering
\begin{tikzpicture}[>=Stealth,semithick]
  \node[kairosstate,initial] (a1) {$Asm_{ok}$};
  \node[kairosstate,right=42mm of a1] (ab) {$Asm_{bad}$};
  \path[->]
    (a1) edge[loop above] node[kairoslabel] {$\ResetOk$} (a1)
    (a1) edge[bend left=16] node[kairoslabel,midway,above] {$\neg \ResetOk$} (ab)
    (ab) edge[loop above] node[kairoslabel] {$\top$} (ab);
\end{tikzpicture}
\caption{Assumption automaton for \texttt{resettable\_delay}.}
\end{figure}

The guarantee automaton keeps track of whether the previous tick was a reset:
\[
Mon_{init},\qquad Mon_{prevR},\qquad Mon_{prevN},\qquad Mon_{bad}.
\]

\begin{figure}[h]
\centering
\begin{tikzpicture}[>=Stealth,semithick]
  \node[kairosstate,initial] (m0) {$Mon_{init}$};
  \node[kairosstate,right=42mm of m0] (mr) {$Mon_{prevR}$};
  \node[kairosstate,below=22mm of mr] (mn) {$Mon_{prevN}$};
  \node[kairosstate,right=45mm of mr] (mb) {$Mon_{bad}$};
  \path[->]
    (m0) edge[bend left=10] node[kairoslabel,midway,above] {$\GuardReset \land reset$} (mr)
    (m0) edge[bend right=10] node[kairoslabel,midway,left] {$\neg reset$} (mn)
    (m0) edge node[kairoslabel,midway,above] {$reset \land \neg \GuardReset$} (mb)
    (mr) edge[loop above] node[kairoslabel] {$\GuardReset \land reset$} (mr)
    (mr) edge[bend left=12] node[kairoslabel,midway,left] {$\GuardAfterReset \land \neg reset$} (mn)
    (mr) edge node[kairoslabel,midway,above]
      {$\neg\bigl((\GuardReset \land reset)\vee(\GuardAfterReset \land \neg reset)\bigr)$} (mb)
    (mn) edge[bend left=12] node[kairoslabel,midway,left] {$\GuardReset \land reset$} (mr)
    (mn) edge[loop below] node[kairoslabel] {$\GuardDelay \land \neg reset$} (mn)
    (mn) edge node[kairoslabel,midway,below]
      {$\neg\bigl((\GuardReset \land reset)\vee(\GuardDelay \land \neg reset)\bigr)$} (mb)
    (mb) edge[loop above] node[kairoslabel] {$\top$} (mb);
\end{tikzpicture}
\caption{Guarantee automaton for \texttt{resettable\_delay}.}
\end{figure}

\paragraph{Pedagogical reading of the example.}
This single example will be reused throughout the paper. At the program level,
\texttt{resettable\_delay} is still small; at the specification level, it combines:
\begin{itemize}[leftmargin=1.4em]
\item a non-trivial input-side safety assumption;
\item a guarantee with distinct reset and non-reset branches;
\item a user invariant that explains the ordinary delayed-data behavior.
\end{itemize}
It is therefore complex enough to show the role of assumptions, guarantees,
user invariants, dangerous steps, generated clauses, and generated triples in a
single coherent running example.

\section{Explicit Product \texorpdfstring{\(program \times A \times G\)}{program x A x G}}

\subsection{Intuition}

The explicit product is not meant to artificially ``glue together'' three
graphs. It represents the following semantic fact: at each tick \(k\),
\begin{enumerate}[leftmargin=1.4em]
\item the program reads \(i(k)\) and produces \(o(k)\);
\item \(A\) reads \(i(k)\);
\item \(G\) reads \((i(k),o(k))\).
\end{enumerate}
There is therefore a \emph{unique synchronized step} of the product.

\subsection{Product States and Steps}

\begin{definition}[Program State and Product State]
We write
\[
\mathsf{PState} := S.
\]
A product state is a triple
\[
\mathsf{ProdState} := S \times Q_A \times Q_G.
\]
\end{definition}

\begin{remark}
This is indeed the \emph{finite} component of the product. Memory, input, and
output do not belong to the product state in the current Rocq formalization;
they appear at the level of the concrete step and the local context.
\end{remark}

\begin{definition}[Product Run]
For an input stream \(u\), we define:
\[
\rps(u,k) :=
\bigl(s_u(k),\, q_A(k),\, q_G(k)\bigr),
\]
where \(q_A\) is the run of \(A\) on \(u\), and \(q_G\) is the run of \(G\) on
\(\run(u)\).
\end{definition}

\begin{definition}[Initial Product State]
The initial product state is
\[
\mathsf{init\_product\_state} := (s_0,q_{A,0},q_{G,0}),
\]
where \(q_{A,0}\) and \(q_{G,0}\) are the initial states of the automata
\(A\) and \(G\).
\end{definition}

\begin{definition}[Product Step]
A product step \(p\) is an object of the form
\[
p = \bigl((s,q_A,q_G),\, m,\, i,\, t,\, e_A,\, e_G \bigr)
\]
such that:
\begin{enumerate}[leftmargin=1.4em]
\item the concrete program step satisfies
  \[
  ((s,m),i)\xrightarrow{\,t\,}_P((s',m'),o);
  \]
\item \(\mathsf{src}_A(e_A)=q_A\) and \(\mathsf{Lab}_A(e_A)(i)=\True\) and
  \(\mathsf{dst}_A(e_A)=q_A'\);
\item \(\mathsf{src}_G(e_G)=q_G\) and \(\mathsf{Lab}_G(e_G)(i,o)=\True\) and
  \(\mathsf{dst}_G(e_G)=q_G'\).
\end{enumerate}
\[
\text{The target of the step is then } (s',q_A',q_G').
\]
\end{definition}

\begin{remark}
The first point of this definition can also be read as
\[
((s,m),i)\xrightarrow{\,t\,}_P((s',m'),o),
\]
which makes the synchronization between the program step, the reading of \(A\)
on \(i\), and the reading of \(G\) on \((i,o)\) more visible, without making
memory part of the finite product state.
\end{remark}

\begin{remark}
Everything is written here as a logical relation:
\[
\StepRel(t,s,m,i,s',m',o),
\qquad
\mathsf{Lab}_A(e_A)(i),
\qquad
\mathsf{Lab}_G(e_G)(i,o).
\]
We deliberately avoid imperative notations such as \(m := i\), which suit the
code but not the mathematical model.
\end{remark}

\subsection{Why This Product Is the Right Object}

The explicit product plays three roles.
\begin{enumerate}[leftmargin=1.4em]
\item It gives a common semantics to the program and the automata.
\item It makes it possible to talk locally about situations in which \(G\) is
  about to be violated.
\item It provides a unique basis for the Rocq proof and for the product IR in
  the current implementation.
\end{enumerate}

In particular, the product explains why a concrete tick is always described by a
minimal quadruple of information:
\[
\text{product source},\qquad \text{input},\qquad \text{output},\qquad \text{product target}.
\]

\subsection{Example: Relevant Subgraph of \texttt{resettable\_delay}}

In \texttt{resettable\_delay}, the most relevant product states are:
\[
\bigl(Init,Asm_{ok},Mon_{init}\bigr),
\qquad
\bigl(Run,Asm_{ok},Mon_{prevR}\bigr),
\qquad
\bigl(Run,Asm_{ok},Mon_{prevN}\bigr),
\qquad
\bigl(Run,Asm_{ok},Mon_{bad}\bigr).
\]
The initial tick leads either to \(Mon_{prevR}\) or to \(Mon_{prevN}\),
depending on the value of \(reset\). Afterwards, two classes of dangerous
transitions matter:
\begin{itemize}[leftmargin=1.4em]
\item reset-labeled dangerous steps, which should be excluded with the help of
  the assumption predicate \(\ResetOk\);
\item non-reset dangerous steps from \(Mon_{prevN}\), which should be excluded
  with the help of the user invariant describing the delayed sample.
\end{itemize}

\begin{figure}[h]
\centering
\begin{tikzpicture}[>=Stealth,semithick]
  \node[kairosstate,initial] (p0) {$(Init,Asm_{ok},Mon_{init})$};
  \node[kairosstate,right=46mm of p0] (pr) {$(Run,Asm_{ok},Mon_{prevR})$};
  \node[kairosstate,below=22mm of pr] (pn) {$(Run,Asm_{ok},Mon_{prevN})$};
  \node[kairosstate,right=48mm of pr] (pb) {$(Run,Asm_{ok},Mon_{bad})$};
  \path[->]
    (p0) edge[bend left=10] node[kairoslabel,midway,above]
      {$\ResetOk \land reset$\\$y = 0$\\$m' = 0$} (pr)
    (p0) edge[bend right=10] node[kairoslabel,midway,left]
      {$\neg reset$\\$y = 0$\\$m' = x$} (pn)
    (pr) edge[bend left=12] node[kairoslabel,midway,left]
      {$\GuardAfterReset \land \neg reset$} (pn)
    (pn) edge[loop below] node[kairoslabel]
      {$\GuardDelay \land \neg reset$\\$y = m$\\$m' = x$} (pn)
    (pr) edge node[kairoslabel,midway,above]
      {$reset \land \neg \GuardReset$} (pb)
    (pn) edge node[kairoslabel,midway,below]
      {$\neg reset \land \neg \GuardDelay$} (pb);
\end{tikzpicture}
\caption{Relevant product subgraph for \texttt{resettable\_delay}.}
\end{figure}

\section{Dangerous Steps and Local Obligations}

\subsection{Dangerous Step}

The central notion of Kairos is now easy to define.

\begin{definition}[Dangerous Step]
A product step
\[
p :
\bigl(s,q_A,q_G\bigr)
\longrightarrow
\bigl(s',q_A',q_G'\bigr)
\]
is \emph{dangerous} if
\[
q_A' \neq bad_A
\qquad\text{and}\qquad
q_G' = bad_G.
\]
\end{definition}

\paragraph{Intuition.}
This is not just any step to a bad state. One must distinguish two cases:
\begin{itemize}[leftmargin=1.4em]
\item if \(A\) itself goes to \(bad_A\), then one leaves the admissible input
  domain; this is not the main objective of the conditional theorem;
\item if \(A\) remains admissible but \(G\) goes to \(bad_G\), then an
  admissible input is enough to produce a guarantee violation. This is the
  situation that must be ruled out locally.
\end{itemize}

\paragraph{Connection to the global semantics.}
If \(\AvoidG(\run(u))\) is false, there exists a first index \(k\) such that
\(q_G(k+1)=bad_G\). Tick \(k\) then realizes a dangerous step. This observation
is what allows one to move from a global violation over an infinite word to a
local witness at a single tick.

\subsection{Local Context and Matching Relation}

\paragraph{Why introduce a local context?}
Obligation generation is static: it ranges over abstract product steps. Actual
execution is dynamic: at each tick it realizes a concrete program transition and
advances \(A\) and \(G\). To connect these two levels, one must summarize the
concrete tick by a local object, and then state when this object realizes a
given abstract step.

\paragraph{Concrete tick versus abstract step.}
The \emph{concrete tick} at rank \(k\) is the step actually observed in an
execution under an input stream \(u\): current configuration, current input,
current output, and next configuration. The \emph{abstract product step} is a
purely static object, enumerated by the generator, which describes a possible
transition of the triple \(program \times A \times G\). Obligations are
generated over abstract steps and then validated over all concrete ticks.

\begin{definition}[Factored Local Context]
We continue to write \(\ctx(u,k)\) for the trace-derived local context
\(\ctx_u(k)\) introduced earlier. Thus,
\[
\ctx(u,k)=\bigl(k,(s_u(k),m_u(k)),u(k),o_u(k),(s_u(k+1),m_u(k+1))\bigr).
\]
The automaton components attached to the same tick are not stored inside
\(\TickCtx\); they are recovered separately from the product run
\(\rps(u,k)=(s_u(k),q_A(k),q_G(k))\) and \(\rps(u,k+1)\).
\end{definition}

\begin{remark}
This matches the underlying Rocq structure exactly: \texttt{StepCtx} carries
only the concrete program tick, while the product-state components are tracked
separately by the product run and by the realized product step.
\end{remark}

\begin{definition}[Matching a Step]
Let
\[
p = \bigl((s,q_A,q_G),m,i,t,e_A,e_G\bigr)
\]
be an abstract product step. We say that a local context \(\ctx\)
\emph{matches the program part of \(p\)}, and we write \(\Match(\ctx,p)\), if:
\begin{enumerate}[leftmargin=1.4em]
\item the current program state stored in \(\ctx\) is \(s\);
\item the current memory stored in \(\ctx\) is \(m\);
\item the current input stored in \(\ctx\) is \(i\);
\item the selected program transition at this tick is \(t\), equivalently
  \(t\) is enabled from \((s,m)\) under \(i\).
\end{enumerate}
\end{definition}

\begin{remark}
This is the primitive matching relation of the Rocq kernel
(\texttt{ctx\_matches\_ps}). The automaton components are handled separately:
\begin{itemize}[leftmargin=1.4em]
\item the source automaton states are recovered from the product run
  \(\rps(u,k)\);
\item the target automaton states are recovered from the well-formedness of
  \(p\) and from the realized program transition.
\end{itemize}
This decomposition is important: the elementary \texttt{NoBad} clause is only a
clause over \(\TickCtx\); the stronger automaton facts used to discharge it are
supplied by helper triples.
\end{remark}

\subsection{Obligation Associated with a Dangerous Step}

\begin{definition}[Elementary Obligation]
To every dangerous product step \(p\), we associate the local formula
\[
\mathsf{obl}_p(\ctx) := \neg \Match(\ctx,p).
\]
\end{definition}

This form explains the exact intuition of the clause:
\begin{quote}
``at the current tick, it is impossible to simultaneously realize the chosen
program transition, the chosen assumption edge, and the guarantee edge leading
to \(bad_G\)''.
\end{quote}

\subsection{Detailed Example: \texttt{resettable\_delay}}

\paragraph{What the generator sees.}
The generator first explores the reachable part of the explicit product. In the
steady phase of \texttt{resettable\_delay}, the two important source states are
\[
(Run,Asm_{ok},Mon_{prevR})
\qquad\text{and}\qquad
(Run,Asm_{ok},Mon_{prevN}).
\]
This already indicates that the guarantee proof splits into two semantic
regimes: the regime immediately following a reset, and the ordinary delayed
data regime.

\paragraph{Two representative dangerous steps.}
The running example has two representative dangerous families.

\emph{Dangerous reset step.}
From \((Run,Asm_{ok},Mon_{prevR})\), consider a step combining:
\begin{enumerate}[leftmargin=1.4em]
\item the reset program transition \(t_{run}^{rst}\), whose relation is
  \[
  y = 0 \land m' = 0;
  \]
\item the assumption edge
  \[
  Asm_{ok} \xrightarrow{\, \ResetOk \,} Asm_{ok};
  \]
\item the dangerous guarantee edge
  \[
  Mon_{prevR} \xrightarrow{\, reset \land \neg \GuardReset \,} Mon_{bad}.
  \]
\end{enumerate}
This dangerous step is relevant because it remains admissible for \(A\), yet
would violate the reset branch of the guarantee.

\emph{Dangerous non-reset step.}
From \((Run,Asm_{ok},Mon_{prevN})\), consider a step combining:
\begin{enumerate}[leftmargin=1.4em]
\item the non-reset program transition \(t_{run}^{keep}\), whose relation is
  \[
  y = m \land m' = x;
  \]
\item the same assumption edge, still satisfied because \(reset\) is false;
\item the dangerous guarantee edge
  \[
  Mon_{prevN} \xrightarrow{\, \neg reset \land \neg \GuardDelay \,} Mon_{bad}.
  \]
\end{enumerate}
This second dangerous family corresponds to the ordinary delayed-data branch of
the specification.

\paragraph{The generated clauses.}
Each dangerous step yields a local clause:
\[
\mathsf{obl}_{rst}(\ctx) := \neg \Match(\ctx,p_{rst}^{bad}),
\qquad
\mathsf{obl}_{keep}(\ctx) := \neg \Match(\ctx,p_{keep}^{bad}).
\]
These clauses only forbid the program part of the two dangerous matchings. The
full contradiction is recovered one level higher, inside the generated
\texttt{NoBad} triples.

\paragraph{How the reset contradiction is obtained.}
For \(p_{rst}^{bad}\), the \texttt{NoBad} triple combines:
\begin{enumerate}[leftmargin=1.4em]
\item the matching fact for the reset transition;
\item the automaton-support fact attached to \((Run,Asm_{ok},Mon_{prevR})\);
\item the assumption-side fact \(\ResetOk\).
\end{enumerate}
The program relation gives \(y=0\), while the assumption support gives
\(x=0\) on reset ticks. The dangerous guarantee edge requires \(y \neq x\).
Hence the triple precondition forces both
\[
y = x
\qquad\text{and}\qquad
y \neq x,
\]
which yields contradiction. This dangerous family is therefore ruled out
primarily by the assumption side.

\paragraph{How the non-reset contradiction is obtained.}
For \(p_{keep}^{bad}\), the \texttt{NoBad} triple uses:
\begin{enumerate}[leftmargin=1.4em]
\item the matching fact for the non-reset transition;
\item the user invariant at state \(Run\);
\item the automaton-support fact attached to \((Run,Asm_{ok},Mon_{prevN})\).
\end{enumerate}
Here the program relation gives
\[
y = m
\qquad\text{and}\qquad
m' = x,
\]
while the invariant implies, in the \(Mon_{prevN}\) regime,
\[
m = prev(x).
\]
The dangerous guarantee edge requires
\[
\neg reset \land \neg \GuardDelay.
\]
Together, these facts force both
\[
y = prev(x)
\qquad\text{and}\qquad
y \neq prev(x),
\]
hence contradiction. This dangerous family is therefore ruled out primarily by
the user-invariant side.

\paragraph{Why this example is better.}
The two dangerous branches play different proof roles:
\begin{itemize}[leftmargin=1.4em]
\item one branch is excluded by assumption support;
\item the other branch is excluded by user-invariant support.
\end{itemize}
This is exactly the kind of interaction that the paper needs to exhibit in a
single running example.

\section{Pseudo-Algorithmic Generation of Clauses and Triples}

\subsection{Why a Procedural Description Is Necessary}

It is not enough to enumerate clause families. To understand the Kairos
architecture, one must explain how the tool constructs them from the product.
The right view is the following:
\begin{enumerate}[leftmargin=1.4em]
\item explore reachable product states;
\item enumerate locally realizable product steps;
\item classify each step;
\item generate, according to this class, the appropriate clauses.
\end{enumerate}

\paragraph{Immediate Instantiation on \texttt{resettable\_delay}.}
This procedure already yields something very concrete:
\begin{enumerate}[leftmargin=1.4em]
\item the relevant reachable states are
  \[
  (Init,Asm_{ok},Mon_{init})
  \quad\text{and}\quad
  (Run,Asm_{ok},Mon_{prevR}),
  \quad
  (Run,Asm_{ok},Mon_{prevN});
  \]
\item from these source states, one enumerates two representative dangerous
  families:
  a reset-dangerous one and a non-reset-dangerous one;
\item both are classified as ``objective no bad'';
\item one adds the local safety clause
  \[
  \mathsf{obl}_{rst}(\ctx) = \neg \Match(\ctx,p_{rst}^{bad})
  \qquad\text{or}\qquad
  \mathsf{obl}_{keep}(\ctx) = \neg \Match(\ctx,p_{keep}^{bad}).
  \]
\end{enumerate}
The general procedural view is therefore not an extra layer: it describes
exactly what the running \texttt{resettable\_delay} case already does.

\subsection{General Scheme}

\paragraph{Input.}
The generator receives:
\begin{itemize}[leftmargin=1.4em]
\item a program \(P\);
\item a specification \(\Phi = (A,G,\UserInv)\).
\end{itemize}

\paragraph{Output.}
It produces a set of semantic clauses, each equipped with:
\begin{itemize}[leftmargin=1.4em]
\item an abstract origin;
\item a state or step of the product;
\item a future backend projection.
\end{itemize}

\subsection{Relational Hoare Triples Built from Generated Clauses}

The core proof does not validate isolated clauses. It validates relational
Hoare triples built from them.

\begin{definition}[Generated Automaton-Support Predicate]
We write
\[
\Support : \mathsf{ProdState} \to (\TickCtx \to \mathsf{Prop})
\]
for the generated helper predicate attached to product states. Intuitively,
\(\Support(x)\) collects the automaton-derived local facts that are later used
to justify transition preconditions while the product is at state \(x\).
\end{definition}

\begin{definition}[Initial Context]
We write
\[
\mathsf{InitCtx}(\ctx) :\Longleftrightarrow
\exists m_0 \in \mathsf{InitMem},\exists u,\; \ctx = \ctx_{m_0,u}(0).
\]
\end{definition}

\begin{definition}[Transition Relation on Local Contexts]
Let \(t \in T\). We define the semantic relation of \(t\) on local contexts by
\[
\mathsf{TransRel}_t(\ctx,\ctx') :\Longleftrightarrow
\exists m_0 \in \mathsf{InitMem},\exists u \in \Stream{I},\exists k \in \Nat,\;
\ctx = \ctx_{m_0,u}(k)
\land
\ctx' = \ctx_{m_0,u}(k+1)
\land
((s_u(k),m_u(k)),u(k)) \xrightarrow{\,t\,}_P
((s_u(k+1),m_u(k+1)),o_u(k)).
\]
\end{definition}

\begin{definition}[Relational Hoare Triple]
A relational Hoare triple is a tuple
\[
ht = (\mathsf{target},\mathsf{Pre},\mathsf{Post},\mathsf{origin},\mathsf{clause}),
\]
where:
\begin{itemize}[leftmargin=1.4em]
\item \(\mathsf{target}\) is either \(\mathsf{Init}\) or \(\mathsf{Step}(t)\)
  for some \(t \in T\);
\item \(\mathsf{Pre},\mathsf{Post} : \TickCtx \to \mathsf{Prop}\);
\item \(\mathsf{origin}\) records the conceptual category of the triple;
\item \(\mathsf{clause}\) is the semantic clause carried by the triple.
\end{itemize}
\end{definition}

\begin{definition}[Validity of a Relational Hoare Triple]
The triple \(ht\) is valid if:
\begin{itemize}[leftmargin=1.4em]
\item if \(\mathsf{target}=\mathsf{Init}\), then
  \[
  \forall \ctx,\; \mathsf{InitCtx}(\ctx) \land \mathsf{Pre}(\ctx)
  \Rightarrow \mathsf{Post}(\ctx);
  \]
\item if \(\mathsf{target}=\mathsf{Step}(t)\), then
  \[
  \forall \ctx,\ctx',\;
  \mathsf{TransRel}_t(\ctx,\ctx') \land \mathsf{Pre}(\ctx)
  \Rightarrow \mathsf{Post}(\ctx').
  \]
\end{itemize}
\end{definition}

\begin{definition}[Canonical Generated Triples]
For a product step \(p\), we write \(t_p\) for its underlying program
transition, \(\mathsf{source}(p)\) for its source product state, and
\(\mathsf{target}(p)\) for its target product state.
If the program-state components of these two product states are \(s\) and
\(s'\), we write them explicitly in the invariant clauses below.

The core generation relation contains five canonical forms of triples:
\begin{enumerate}[leftmargin=1.4em]
\item an initial triple for the user invariant:
  \[
  \mathsf{InitInvTriple} :
  \mathsf{Init} \;\{ \top \}\; \{ \UserInv(s_0) \};
  \]
\item an initial triple for automaton support:
  \[
  \mathsf{InitSuppTriple} :
  \mathsf{Init} \;\{ \top \}\; \{ \Support(\mathsf{init\_product\_state}) \};
  \]
\item for each well-formed product step \(p\), a propagation triple for the
  user invariant:
  \[
  \mathsf{InvTriple}_p :
  \mathsf{Step}(t_p)\;
  \{ \Match(\cdot,p) \land \UserInv(s) \}\;
  \{ \UserInv(s') \};
  \]
\item for each well-formed product step \(p\), a propagation triple for
  automaton support:
  \[
  \mathsf{SuppTriple}_p :
  \mathsf{Step}(t_p)\;
  \{ \Match(\cdot,p) \land \Support(\mathsf{source}(p)) \}\;
  \{ \Support(\mathsf{target}(p)) \};
  \]
\item for each well-formed product step \(p\), a safety triple:
  \[
  \mathsf{NoBadTriple}_p :
  \mathsf{Step}(t_p)\;
  \{ \Match(\cdot,p) \land \UserInv(s) \land \Support(\mathsf{source}(p)) \}\;
  \{ \False \}.
  \]
\end{enumerate}
\noindent
This last triple matters semantically when \(p\) is dangerous: in that case, a
valid \(\mathsf{NoBadTriple}_p\) rules out the concrete realization of \(p\).
\end{definition}

\begin{remark}
The paper writes these triples schematically. In Rocq, their exact definitions
are \texttt{init\_node\_inv\_triple},
\texttt{init\_support\_automaton\_triple},
\texttt{node\_inv\_triple ps},
\texttt{support\_automaton\_triple ps}, and
\texttt{no\_bad\_triple ps}. The point of the present presentation is to make
their logical role explicit before discussing backend grouping.
\end{remark}

\begin{definition}[Generated Clauses and Generated Triples]
We write \(\GeneratedBy(o,c)\) when the semantic clause \(c\) is generated with
origin \(o\), and \(\Generated(c)\) when such an origin exists. Concretely, the
core generation relation contains:
\begin{itemize}[leftmargin=1.4em]
\item the two initial clauses, for node invariants and automaton support;
\item for each well-formed product step \(p\), one user-invariant propagation
  clause, one automaton-support propagation clause, and one objective clause
  \(\mathsf{NoBad}_p(\ctx)=\neg \Match(\ctx,p)\).
\end{itemize}
Similarly, \(\GeneratedTriple(ht)\) means that \(ht\) is one of the five
canonical triples listed above, generated either from the initial state or from
a well-formed product step.
\end{definition}

\subsection{Conceptual Algorithm}

\begin{quote}
\textbf{For each} reachable product state \(x=(s,q_A,q_G)\):

\quad 1. Add the initial/coherence clauses attached to \(x\).

\quad 2. Add the user-invariant propagation clauses attached to \(x\), for example
\[
\mathsf{InvObl}_x(\ctx) :=
\bigl(\mathsf{src\_state}(\ctx)=s \land \mathsf{src\_a}(\ctx)=q_A \land \mathsf{src\_g}(\ctx)=q_G\bigr)
\Rightarrow
\bigl(\UserInv(s)(\ctx)\Rightarrow \mathsf{Shift}(\UserInv(s'))(\ctx)\bigr).
\]

\quad 3. Add the automaton-support clauses attached to
\quad \;\;\; \(x\), that is, the formulas that justify the additional local
\quad \;\;\; automaton facts assumed by the backend.

\quad 4. \textbf{For each} product step \(p\) outgoing from \(x\):

\qquad a. If the target of \(p\) falls into \(bad_A\), only add the support
\qquad\;\;\; clauses that are still required.

\qquad b. If \(p\) is dangerous, add the objective clause
\[
\mathsf{ObjObl}_p(\ctx) := \neg \Match(\ctx,p).
\]

\qquad c. If \(p\) is safe, only add the coherence and support clauses
\qquad\;\;\; required for this step.
\end{quote}

\noindent
The point of this procedural presentation is that \emph{NoBad clauses are not
expected to be discharged in isolation}. The backend proves them inside
transition-level Hoare bundles, under extra local assumptions, and those
assumptions are themselves justified by generated helper clauses. The local
proof architecture therefore has one safety layer and one helper layer:
\begin{enumerate}[leftmargin=1.4em]
\item \(\mathsf{Helper/InitGoal}\): establish, at the initial tick, both the
  user-invariant facts and the automaton-support facts;
\item \(\mathsf{Helper/Propagation}\): propagate, from one tick to the next,
  both the user-invariant facts and the automaton-support facts;
\item \(\mathsf{Safety/NoBad}\): exclude dangerous product steps under this
  helper context.
\end{enumerate}

In the current Rocq core, this architecture is reconstructed in two steps.
First, the concrete generation relation produces semantic clauses. Second,
these clauses are assembled into relational Hoare triples whose validity is
stated against the relational semantics of program transitions. The concrete
generation relation contains:
\[
\mathsf{prod\_clause}(ps)(\ctx) := \neg \ctx\mathsf{\_matches\_ps}(\ctx,ps)
\]
for the objective layer, a preservation-style clause for node invariants, and a
preservation-style clause for automaton support. The Rocq kernel now turns such
clauses into relational Hoare triples; the implementation further groups these
triples by program transition before translating them to Why3-facing
verification tasks. What is still more fine-grained on the
implementation side is the internal distinction between monitor-compatibility
support and state-aware assumption support.

The implementation now exposes both levels explicitly:
\begin{itemize}[leftmargin=1.4em]
\item a four-way conceptual projection used to summarize generated clauses as
  \(\mathsf{NoBad}\), \(\mathsf{InitialGoal}\),
  \(\mathsf{UserInvariant}\), and \(\mathsf{AutomatonSupport}\);
\item a finer backend taxonomy that keeps separate OBC/Why3-facing families
  such as \texttt{FamNoBadRequires}, \texttt{FamNoBadEnsures},
  \texttt{FamInitialCoherencyGoal}, \texttt{FamCoherencyRequires},
  \texttt{FamCoherencyEnsuresShifted},
  \texttt{FamMonitorCompatibilityRequires}, and
  \texttt{FamStateAwareAssumptionRequires}.
\end{itemize}
The finer families are projected onto the same four conceptual categories as
the Rocq proof architecture. These four categories are a classification of
\emph{generated clauses}, not four independent proof phases.

\subsection{Four Clause Categories}

The generated semantic clauses fall into four conceptual categories.

\paragraph{NoBad.}
These are the genuine safety objectives. For a dangerous product step \(p\), the
semantic clause is:
\[
\mathsf{NoBad}_p(\ctx) := \neg \Match(\ctx,p).
\]
The backend does not normally send this clause in isolation to the external
tool. The kernel first turns it into a relational Hoare triple whose
postcondition is \(\bot\) on the dangerous branch; the implementation then
places that triple inside a transition-level proof bundle together with the
helper preconditions and postconditions needed to make it provable.
These clauses are sufficient to conclude that the product never reaches
\(bad_G\), but they are not expected to be discharged without helper clauses.

\paragraph{InitialGoal.}
These clauses provide the base case of the induction. They establish that
the helper context already holds at tick \(0\). Concretely, this includes:
\begin{itemize}[leftmargin=1.4em]
\item initialization of the user invariant attached to the initial node;
\item initialization of the automaton-support facts attached to the initial
  product state.
\end{itemize}
In the implementation, they correspond to explicit initial helper triples grouped
in the same initial helper bundle. They could equivalently be encoded by a
ghost predecessor carrying \(\top\), but keeping them explicit is simpler and
clearer.

\paragraph{UserInvariant.}
These clauses justify the node invariants used as local preconditions. The
implementation injects:
\begin{itemize}[leftmargin=1.4em]
\item \(\mathsf{Inv}(\mathsf{src})\) in transition preconditions;
\item \(\mathsf{Shift}(\mathsf{Inv}(\mathsf{dst}))\) in transition
  postconditions.
\end{itemize}
So the predecessor proves the fact that the successor later consumes as
its own precondition. In the Rocq kernel this is expressed by a relational
Hoare triple whose postcondition talks about the successor context. This is one
\emph{subfamily} of the common helper propagation layer. These clauses must be
bundled together with automaton
support clauses for the same transition, because each side may help prove the
other.

\paragraph{AutomatonSupport.}
These clauses justify the automaton-derived facts added to transition
preconditions. There are two important subcases in the implementation:
\begin{itemize}[leftmargin=1.4em]
\item monitor compatibility clauses: if the current monitor state is \(q\),
  then one of the incoming guards leading to \(q\) must have held at the
  previous tick, after shifting;
\item state-aware assumption/product support clauses: for a fixed program
  transition and current guarantee state, one of the locally compatible
  \((A\text{-guard} \land G\text{-guard})\) disjuncts must hold.
\end{itemize}
This is the second \emph{subfamily} of the common helper propagation layer.
Again, in the Rocq kernel these clauses feed relational Hoare triples whose
postconditions speak about the successor context. They are not meant to live in
proof objects separate from user
invariants: they must be discharged in the same helper bundles.

Transition-level user contracts remain visible in the backend as separate
\texttt{transition\_requires} and \texttt{transition\_ensures} families, but
they do not belong to the four generated-proof categories above. They are part
of the user specification surface, not clauses generated by the local
product proof architecture itself.

\paragraph{Why the helper/safety split is needed.}
The soundness point is the following:
\begin{quote}
it is valid to enrich a local dangerous-step proof with additional
preconditions only if generated clauses separately establish that these
preconditions do hold on concrete executions.
\end{quote}
Therefore:
\begin{enumerate}[leftmargin=1.4em]
\item \(\mathsf{Helper/InitGoal}\) provides the base case for both invariant
  facts and automaton-support facts;
\item \(\mathsf{Helper/Propagation}\) propagates both kinds of helper facts
  through time, inside the same transition-level proof objects;
\item \(\mathsf{Safety/NoBad}\) excludes the dangerous product steps under this
  common helper context.
\end{enumerate}

\begin{remark}
The four labels
\(\mathsf{NoBad}\), \(\mathsf{InitialGoal}\), \(\mathsf{UserInvariant}\), and
\(\mathsf{AutomatonSupport}\) are useful as a classification of generated
clauses. However, the proof architecture is not a chain of four isolated
layers: \(\mathsf{UserInvariant}\) and \(\mathsf{AutomatonSupport}\) must be
bundled together inside the same helper proof objects, both at initialization
and at propagation time.
\end{remark}

\section{Proof Structure}

\subsection{Global Assumptions of the Core Theorem}

The final theorem of the core does not model an external checker. It is stated
directly under two groups of hypotheses.

\paragraph{Program-model well-formedness.}
The program must satisfy the structural assumptions captured by the Rocq record
\texttt{WellFormedProgramModel}:
\begin{itemize}[leftmargin=1.4em]
\item the selected program transition is enabled at every tick;
\item the selected edges of \(A\) and \(G\) do start from the current automaton
  states;
\item the labels of these selected edges are true of the current input and
  current observation;
\item the initial states of \(A\) and \(G\) are not bad.
\end{itemize}
These assumptions express that the semantic model is total on the class of
programs under consideration. They are complemented by the separate initial
memory admissibility predicate \(\mathsf{InitMem}\), which characterizes the class of
initial memories on which the theorem is stated.

\paragraph{Validation hypothesis.}
The second global hypothesis states that all generated relational Hoare triples
are valid:
\[
\forall ht,\; GeneratedTriple(ht) \Rightarrow TripleValid(ht).
\]
Everything else is proved inside the core from this hypothesis.

\subsection{Step 1: Product Progression}

\begin{proposition}[Progress of the Explicit Product]
For every admissible initial memory \(m_0 \in \mathsf{InitMem}\), every input stream
\(u\), and every tick \(k\), there exists a well-formed product step \(ps\)
such that:
\begin{enumerate}[leftmargin=1.4em]
\item \(ps\) is the step realized at tick \(k\);
\item the target of \(ps\) is exactly the product state reached at tick \(k+1\).
\end{enumerate}
\end{proposition}

\begin{proof}
This is theorem \texttt{product\_progresses\_at\_each\_tick} in the Rocq
development. The proof is by explicit construction: the semantic selectors of
the program, of \(A\), and of \(G\) define a canonical step
\(\mathsf{product\_select\_at}(u,k)\). The well-formedness part follows from
\texttt{WellFormedProgramModel}. The target computation is then checked by
unfolding the definitions of the program step, of the automaton updates, and of
the local observation sent to \(G\).
\end{proof}

\begin{proposition}[A Realized Step Induces the Transition Relation]
Let \(u\) be an input stream, \(k\) a tick, and \(p\) a product step realized
at tick \(k\). Then the underlying program transition \(t_p\) satisfies
\[
\mathsf{TransRel}_{t_p}\bigl(\ctx(u,k),\ctx(u,k+1)\bigr).
\]
\end{proposition}

\begin{proof}
This is the content of the Rocq lemma
\texttt{transition\_rel\_of\_realized\_step}. Once \(p\) is known to be the
step realized at tick \(k\), the relation \(\mathsf{TransRel}_{t_p}\) follows
immediately from its definition.
\end{proof}

\subsection{Step 2: From a Global Violation to a Dangerous Local Step}

\begin{proposition}[Global Violation Implies a Dangerous Tick]
For every admissible initial memory \(m_0 \in \mathsf{InitMem}\) and every input stream
\(u\), if \(\AvoidA(u)\) and \(\neg \AvoidG(\run_{m_0}(u))\), then there exists
a tick \(k\) such that \(bad\_local\_step(m_0,u,k)\).
\end{proposition}

\begin{proof}
This is theorem \texttt{bad\_local\_step\_if\_G\_violated}. Since
\(\neg \AvoidG(\run(u))\), there exists a first index \(j\) such that
\(q_G(j)=bad_G\). The initial state of \(G\) is not bad, hence \(j=S\,k\) for
some \(k\). By Step~1, a concrete well-formed product step is realized at tick
\(k\); its target is the product state reached at tick \(k+1\). The guarantee
component of this target is therefore \(bad_G\). Because \(\AvoidA(u)\) holds,
the assumption component of the same target is not \(bad_A\). This is exactly
the definition of a dangerous local step.
\end{proof}

\subsection{Step 3: Helper Facts Hold on Admissible Runs}

\begin{proposition}[Initial Helper Facts]
For every admissible initial memory \(m_0 \in \mathsf{InitMem}\) and every admissible
run on \(u\), the two helper facts needed by the local proof are true at the
initial tick:
\begin{enumerate}[leftmargin=1.4em]
\item the user invariant attached to the initial program state holds;
\item the automaton-support fact attached to the initial product state holds.
\end{enumerate}
\end{proposition}

\begin{proof}
These are propositions \texttt{init\_node\_inv\_holds} and
\texttt{init\_support\_automaton\_holds}. Each one is obtained by applying the
validity hypothesis to a generated initial triple:
\texttt{init\_node\_inv\_triple} or
\texttt{init\_support\_automaton\_triple}. The precondition of these triples is
the initial-context predicate \texttt{init\_ctx}, which holds by construction of
\(\ctx_{m_0,u}(0)\). Their postconditions are precisely the helper facts to be
established at tick \(0\).
\end{proof}

\begin{proposition}[Propagation of Helper Facts]
Let \(m_0 \in \mathsf{InitMem}\) and let \(u\) be an admissible input stream. Then for
every tick \(k\):
\begin{enumerate}[leftmargin=1.4em]
\item the user invariant of the current program state holds on \(\ctx(u,k)\);
\item the automaton-support fact of the current product state holds on
  \(\ctx(u,k)\).
\end{enumerate}
\end{proposition}

\begin{proof}
This is the content of propositions \texttt{node\_inv\_holds\_on\_run} and
\texttt{support\_automaton\_holds\_on\_run}. Both proofs are inductions on
\(k\).

The base case is given by the previous proposition. For the induction step, the
canonical realized product step at tick \(k\) yields the generated propagation
triples \texttt{node\_inv\_triple ps} and
\texttt{support\_automaton\_triple ps}. By the induction hypothesis, the
corresponding preconditions hold on \(\ctx(u,k)\). By validity of generated
triples, together with the transition-relation proposition of Step~1, their
postconditions hold on \(\ctx(u,k+1)\). This establishes the two helper facts
at the successor tick.
\end{proof}

\begin{proposition}[Combined Helper Context]
Let \(m_0 \in \mathsf{InitMem}\) and let \(u\) be admissible. If a well-formed product
step \(ps\) is realized at tick \(k\), then the full helper context required by
the safety triple attached to \(ps\) holds on \(\ctx_{m_0,u}(k)\).
\end{proposition}

\begin{proof}
This is proposition \texttt{helper\_context\_holds\_on\_run}. It is an immediate
combination of the two propagation results above, instantiated at the source
program state and source product state of the realized step \(ps\).
\end{proof}

\subsection{Step 4: A Dangerous Tick Produces a Falsified Generated Clause}

\begin{proposition}[Clause Coverage of Dangerous Steps]
For every admissible initial memory \(m_0 \in \mathsf{InitMem}\), every input stream
\(u\), and every tick \(k\), if \(bad\_local\_step(m_0,u,k)\), then there
exists a generated semantic clause \(c\) such that:
\[
Generated(c)
\qquad\text{and}\qquad
\neg c(\ctx_{m_0,u}(k)).
\]
\end{proposition}

\begin{proof}
This is theorem \texttt{generation\_coverage}. By definition of
\(bad\_local\_step\), a dangerous tick comes with a well-formed realized product
step \(ps\). The core associates to \(ps\) the safety clause
\texttt{prod\_clause ps}, whose meaning is that the current concrete tick does
not match this dangerous step. But a local lemma
\texttt{ctx\_at\_matches\_realized\_ps} proves that the concrete context
\(\ctx(u,k)\) does match the realized step \(ps\). Therefore the clause
\texttt{prod\_clause ps} is falsified at \(\ctx(u,k)\).
\end{proof}

\subsection{Step 5: A Dangerous Tick Produces an Applicable \texttt{NoBad} Triple}

\begin{proposition}[Triple Coverage of Dangerous Steps]
For every admissible initial memory \(m_0 \in \mathsf{InitMem}\), every input stream
\(u\), and every tick \(k\), if \(bad\_local\_step(m_0,u,k)\), then there
exist a product step \(ps\) and a generated triple \(ht\) such that:
\begin{enumerate}[leftmargin=1.4em]
\item \(ht = \texttt{no\_bad\_triple}\;ps\);
\item the precondition of \(ht\) holds on \(\ctx(u,k)\);
\item the postcondition of \(ht\) is \(\False\).
\end{enumerate}
\end{proposition}

\begin{proof}
This is theorem \texttt{triple\_generation\_coverage}. Again, a dangerous tick
provides a realized dangerous product step \(ps\). The core generates from this
step the triple \texttt{no\_bad\_triple ps}. Its postcondition is by
construction the always-false clause \texttt{FalseClause}. Its precondition is
the conjunction of three ingredients:
\begin{enumerate}[leftmargin=1.4em]
\item the matching condition saying that the current tick realizes \(ps\);
\item the user-invariant fact required at the current source state;
\item the automaton-support fact required at the current source product state.
\end{enumerate}
The first item comes from \texttt{ctx\_at\_matches\_realized\_ps}; the other two
come from the combined helper-context corollary of Step~3. Hence the
precondition of \texttt{no\_bad\_triple ps} does hold on the dangerous tick.
\end{proof}

\subsection{Step 6: Valid Triples Rule Out Global Violations}

\begin{theorem}[Conditional Correctness]
Assume \texttt{WellFormedProgramModel} and assume that every generated triple is
valid. Then, for every admissible initial memory \(m_0 \in \mathsf{InitMem}\) and every
input stream \(u\),
\[
\AvoidA(u) \Rightarrow \AvoidG(\run_{m_0}(u)).
\]
\end{theorem}

\begin{proof}
This is theorem \texttt{triple\_valid\_conditional\_correctness\_under\_wf} in
the Rocq development. Fix an admissible initial memory \(m_0 \in \mathsf{InitMem}\), an
input stream \(u\), and assume \(\AvoidA(u)\). Suppose, for contradiction,
that \(\neg \AvoidG(\run_{m_0}(u))\). By Step~2, there exists a dangerous tick
\(k\). By Step~5, there exist a dangerous realized
product step \(ps\) and a generated triple
\(ht = \texttt{no\_bad\_triple}\;ps\) such that:
\begin{enumerate}[leftmargin=1.4em]
\item the precondition of \(ht\) holds on \(\ctx(u,k)\);
\item the postcondition of \(ht\) is \(\False\);
\item the transition of \(ps\) is exactly the transition realized at tick \(k\).
\end{enumerate}
Since \(ht\) is generated, the global validity hypothesis yields
\(TripleValid(ht)\). Instantiating this validity statement with the concrete
tick \(k\) and the concrete successor context \(\ctx(u,k+1)\), one obtains that
the postcondition of \(ht\) must hold at the successor tick; Step~1 provides
the required transition-relation premise. But that
postcondition is \(\False\), a contradiction. Therefore no global violation of
\(G\) can occur on an admissible input stream, and \(\AvoidG(\run(u))\) follows.
\end{proof}

\begin{remark}
The clause-level theorem \texttt{generation\_coverage} and the triple-level
theorem \texttt{triple\_generation\_coverage} play different roles. The former
is the sharp semantic reduction from a dangerous local step to a falsified
generated clause. The latter is the bridge actually used by the final proof: it
shows that the same dangerous local step also yields a generated validatable
object, namely a \texttt{NoBad} relational Hoare triple whose precondition is
satisfied on the concrete tick.
\end{remark}

\section{Implementation Language and Why3 Backend}

\subsection{Source Language of the Implementation}

The mathematical model introduced above is reflected in the current Kairos
implementation through two layers that are kept conceptually distinct.

\paragraph{Program layer.}
A node contains:
\begin{itemize}[leftmargin=1.4em]
\item input and output declarations;
\item local variables and state variables;
\item a finite set of control states and an initial control state;
\item a transition relation described by guarded node transitions.
\end{itemize}
This is the executable part of the source program.

\paragraph{Specification layer.}
The same node carries:
\begin{itemize}[leftmargin=1.4em]
\item assumptions, interpreted over input traces;
\item guarantees, interpreted over reactive input/output traces;
\item state-related user invariants, interpreted over local trace contexts.
\end{itemize}
In the implementation, this split is explicit: the program part provides the
transition semantics, while the specification part provides the formulas from
which automata and support clauses are derived.

\subsection{From Detailed Clauses to Backend Code}

The detailed semantic clauses defined in this paper are not sent to Why3 as raw
\(\neg \Match(\ctx,p)\) formulas over abstract product steps. They are first
assembled into relational Hoare triples and then projected to a finite backend
representation.

\paragraph{Step 1: product analysis.}
The implementation builds or explores the product
\(program \times A \times G\). This exploration computes:
\begin{itemize}[leftmargin=1.4em]
\item reachable product states;
\item realizable product steps;
\item the classification of steps into safe, dangerous, inadmissible, or
  pruned;
\item the abstract local clauses attached to those states and steps.
\end{itemize}

\paragraph{Step 2: backend projection.}
These abstract clauses and triples are then compiled into an instrumented
program:
\begin{itemize}[leftmargin=1.4em]
\item automaton information is represented by finite backend variables such as
  \texttt{\_\_aut\_state};
\item references to past values are compiled through auxiliary variables such
  as \texttt{\_\_pre\_k\ldots};
\item local clauses become backend preconditions, postconditions, linking
  invariants, and transition-side formulas.
\end{itemize}
The purpose of this projection is not to change the semantics, but to replace
the explicit matching relation over product steps with formulas that Why3 can
attach to concrete program declarations.

\paragraph{Step 3: Why3 generation.}
From this instrumented intermediate form, Kairos produces:
\begin{itemize}[leftmargin=1.4em]
\item an OBC-like textual form used for debugging and inspection;
\item a Why3 abstract syntax tree;
\item a rendered Why3 source file;
\item optional verification-condition and SMT dumps, one block per goal.
\end{itemize}
At this stage, the triples of the paper have become Why3 verification
conditions attached to functions, transitions, and auxiliary lemmas of the
generated program~\cite{why3}.

\subsection{Concrete Walkthrough on \texttt{resettable\_delay}}

The current toolchain can already be exercised on a concrete Kairos source
version of the running example. Since this section is also meant to illustrate
the concrete syntax of the implementation language, we reproduce the full node:
\begin{verbatim}
node resettable_delay(reset: int, x: int) returns (y: int)
contracts
  requires: G((reset = 0 or reset = 1) and (reset = 1 => x = 0));
  ensures: G((reset = 1) => (y = 0));
  ensures: X G((reset = 0 and prev reset = 1) => y = 0);
  ensures: X G((reset = 0 and prev reset = 0) => y = prev x);
locals
  m : int;
states
  Init(init), Run;
invariants
  in Run:
    (prev reset = 1 => m = 0) and
    (prev reset = 0 => m = prev x);
transitions
  Init:
    to Run when reset = 1 {
      y := 0;
      m := 0;
    }
    to Run when reset = 0 {
      y := 0;
      m := x;
    }
  Run:
    to Run when reset = 1 {
      y := 0;
      m := 0;
    }
    to Run when reset = 0 {
      y := m;
      m := x;
    }
end
\end{verbatim}
The actual source used in the implementation repository is
\texttt{tests/ok/inputs/resettable\_delay.kairos}; the generated artifacts used
below are produced directly from that file with the V2 CLI.

\paragraph{Residual automata produced by the tool.}
On this example, the current implementation prints the following residual
states:
\begin{itemize}[leftmargin=1.4em]
\item assumption side: \(A0\) and \(A1\), where \(A1\) is the inconsistent
  residual \(\False\);
\item guarantee side: \(G0\), \(G1\), and \(G2\), where \(G1\) is the
  inconsistent residual \(\False\) and \(G2\) is the steady-state residual
  obtained after the first tick.
\end{itemize}
This is the concrete manifestation of the usual safety-monitoring workflow:
state predicates are first propositionalized into tick-local atoms, then
compiled into residual temporal states, and finally rendered as finite safety
automata.

\paragraph{Reachable product and generated clauses.}
The current product exploration reports five reachable states:
\[
(Init,A0,G0),\quad
(Run,A0,G1),\quad
(Run,A0,G2),\quad
(Run,A1,G1),\quad
(Run,A1,G2).
\]
For this concrete node, Kairos also reports \(22\) generated clauses, split as
follows:
\begin{itemize}[leftmargin=1.4em]
\item \(3\) safety clauses;
\item \(19\) helper clauses;
\item among helper clauses, \(1\) initial goal and \(18\) propagation clauses;
\item among generated clause kinds, \(6\) come from user invariants and \(12\)
  from automaton support.
\end{itemize}
This is already enough to illustrate the architecture of the paper on a real
example: the source specification is compiled to residual automata, the product
is explored, and the resulting dangerous or supportive situations are translated
to local proof objects.

\paragraph{Instrumented OBC and generated Why3 code.}
The generated OBC-like form makes the backend projection explicit. For
\texttt{resettable\_delay}, the tool introduces the ghost fields
\texttt{\_\_aut\_state}, \texttt{\_\_pre\_k1\_x}, and
\texttt{\_\_pre\_k1\_reset}. Transition contracts then mention:
\begin{itemize}[leftmargin=1.4em]
\item compatibility facts about the monitor state;
\item assumption-side support formulas;
\item coherence formulas carrying the user invariant;
\item safety formulas excluding \texttt{NoBad} combinations.
\end{itemize}
For instance, the concrete extracted OBC-like transition for the initial reset
case contains the expected combination of compatibility, assumption-side
support, a \texttt{NoBad} postcondition, and the shifted coherence fact:
\begin{verbatim}
transition Init -> Run when reset = 1 {
  require __aut_state = Aut0                          {Compatibility};
  require __aut_state = Aut0 -> ...                   {AssumeAutomaton};
  ensure  __aut_state = Aut0 -> not (...)             {Instrumentation};
  ensure  (reset = 1 -> m = 0) and (reset = 0 -> m = x) {Coherency};
  do
    y := 0;
    m := 0;
}
\end{verbatim}
The corresponding Why3 file contains the same information in Why3 syntax: an
enumerated type for monitor states, a record for program variables including
ghost fields, a \texttt{step} procedure, and the projected preconditions and
postconditions.

\paragraph{Validation status on the current backend.}
The extraction pipeline succeeds end-to-end on this example: automata,
reachable product, clause map, OBC-like code, and Why3 source are all
generated. On the current backend, the resulting Why3 development is discharged
by Z3 when Why3 is run with the standard simplification and splitting
transformations used for transition VCs:
\begin{verbatim}
tmpconf=$(mktemp /tmp/why3conf.XXXXXX)
why3 config --config "$tmpconf" detect
why3 -C "$tmpconf" prove \
  -a simplify_formula \
  -a eliminate_if_term \
  -a remove_unused \
  -a split_vc \
  -P z3 -t 30 resettable_delay.mlw
\end{verbatim}
In our current environment, this command exits successfully on the generated
file. This matters for the paper for two reasons:
\begin{itemize}[leftmargin=1.4em]
\item it confirms that the mathematical reduction can already be exercised on a
  concrete Kairos example through the implementation path;
\item it makes visible the exact boundary between the semantic proof objects of
  the paper and the backend proof obligations actually discharged by Why3.
\end{itemize}

\subsection{What Is Actually Sent to Why3}

The object sent to Why3 is therefore not the mathematical product itself, but a
finite program annotated with the formulas that result from the projection
above. Concretely, the generated Why3 file contains:
\begin{itemize}[leftmargin=1.4em]
\item declarations for program states, memories, and auxiliary variables;
\item one or more Why3 procedures encoding the program step relation;
\item preconditions and postconditions representing the projected clauses;
\item additional formulas linking source variables, target variables, automaton
  states, and history variables.
\end{itemize}

From the point of view of the present paper, the essential invariant is the
following: every backend verification condition must correspond to a semantic
fact justified by the product analysis. In particular, dangerous product steps
must give rise to formulas whose failure would exactly witness the local
realization of such a step.

\paragraph{Additional structural side conditions.}
Because the abstract model assumes that every tick computes a successor state,
a successor memory, and an output valuation, a practical Kairos backend may
need companion checks or side conditions ensuring that the concrete program
falls within this fragment. Such conditions concern, for instance, the absence
of runtime blocking caused by invalid operations, the fact that each output
signal is assigned a value at every instant, and the exhaustiveness of the
transition filtering by current state and memory. These conditions do not
belong to the abstract product core itself; they are prerequisites for applying
the abstract semantics to the concrete source language.

\subsection{What Kairos Retrieves from Why3}

When Why3 is invoked, Kairos receives a collection of proof goals together with
their statuses. For each goal, the relevant result is one of the standard Why3
outcomes: proved, invalid, unknown, timeout, or backend failure.

\paragraph{Meaning of a positive result.}
If a goal is proved, then the corresponding projected backend formula is valid
in the logic handled by Why3 and its external provers. Under the intended
correctness theorem of the paper, such a result is used as evidence that the
associated local triple is valid, and therefore that its underlying clause or
support condition holds. This also applies to
any side conditions or companion checks used to guarantee that the concrete
program belongs to the total fragment assumed by the model.

\paragraph{Meaning of a negative or inconclusive result.}
If a goal is reported invalid, unknown, timed out, or failed, the mathematical
meaning is weaker than ``the property is false''. It means that Kairos did not
obtain external validation for the projected formula. This may reveal:
\begin{itemize}[leftmargin=1.4em]
\item a real bug in the program or in the specification;
\item an insufficient invariant or support clause;
\item a weakness of the backend encoding;
\item or a limitation of the automatic prover.
\end{itemize}
Hence the result returned by Why3 must always be interpreted relative to the
projection correctness assumptions of the overall method.

\subsection{Why the Backend Still Matters}

The explicit product remains the right semantic reference, but it is not the
right final object to materialize exhaustively in generated Why3 code. The
backend remains necessary in order to:
\begin{itemize}[leftmargin=1.4em]
\item keep the generated code finite and manageable;
\item encode automaton and history information through ordinary program
  variables;
\item expose triples and their projected clauses in a format accepted by Why3 and SMT solvers.
\end{itemize}
The methodological split is therefore the following:
\begin{itemize}[leftmargin=1.4em]
\item the product provides the semantic source of the clauses;
\item the backend packages them into transition-level triples and provides
  their finite proof-oriented encoding.
\end{itemize}

\section{Related Work}

The closest related work falls into three technical families. The first one
reduces temporal properties of programs to local proof material checked by a
deductive backend. The second one studies synchronous contracts, observers, and
assume/guarantee reasoning over reactive systems. The third one compiles
temporal properties into executable monitoring artifacts. Kairos borrows an
essential idea from each family, but differs from all of them in the semantic
role assigned to the explicit product \(program \times A \times G\). This
comparison matters because the contribution of Kairos is not merely another
automata-based verifier: its main claim is that the explicit product is the
right semantic object for transporting a global conditional safety property
toward local deductive proof objects over a program that may manipulate
unbounded data.

For clarity, the comparison should be read along two axes. The first axis is
\emph{what is proved}: a property of program traces, a contract over a
synchronous transition system, or a runtime-monitoring obligation. The second
axis is \emph{how the proof is mediated}: through annotations attached to
program points, through invariants over a synchronous transition system, or
through an explicit semantic product between a program and temporal automata.
Kairos is best understood as occupying the corner ``reactive program theorem +
explicit product + deductive local proof objects''.

Another useful axis of comparison is finite-state versus unbounded-state
reasoning. Classical temporal model checking is most natural when the system
and the temporal monitor can be explored in one explicit transition system. The
present setting is different: the automata are finite-state, but the program
state space need not be. Kairos therefore combines finite safety automata with
deductive transition reasoning instead of reducing the whole problem to
explicit-state exploration. The contribution is thus not an explicit-state
algorithm for a finite product, but a proof architecture in which the finite
temporal component and the deductive program component meet at the level of
product steps, generated clauses, and relational Hoare triples.

\subsection{From Temporal Properties to Program Proof Objects}

\paragraph{Aorai.}
Aorai~\cite{aorai-manual,stouls-groslambert-2009} is probably the closest
neighbor to Kairos if one focuses on the idea of compiling temporal properties
into local proof material attached to a program. In Aorai, LTL properties about
C programs are translated into automata and then into ACSL-style annotations
checked by the Frama-C toolchain. The key similarity with Kairos is the
reduction from a global temporal property to finitely many local proof
objects. The main difference is semantic. Aorai is centered on annotated C
programs and temporal properties over program traces, whereas Kairos is
centered on reactive traces with an explicit distinction between input
assumptions and input/output guarantees. Kairos also makes the product
\(program \times A \times G\) explicit as a semantic object, instead of keeping
the automaton/program synchronization mostly implicit in the generated
annotations. This explicit product is not cosmetic: it is what lets Kairos
state and prove the reduction from a global guarantee violation to a dangerous
local step, which is the central proof idea of the present work.

\paragraph{CaFE.}
CaFE~\cite{cafe-page} extends the same general ecosystem to CaRet properties,
which are tailored to nested calls and returns. This makes it especially
relevant when temporal reasoning must account for call-stack structure. Its
proximity to Kairos lies in the idea that temporal reasoning over programs can
be reduced to local proof objects checked by an existing backend. Its
difference from Kairos lies in the observed semantic domain: CaFE is organized
around call/return structure of C executions, whereas Kairos is organized
around reactive ticks, stateful stream semantics, and assume/guarantee safety on
input/output traces. In particular, Kairos is not merely ``Aorai for
synchronous programs'': its target property is conditional safety under input
admissibility, and its core semantic object already combines program and
automata state.

\subsection{Assume/Guarantee and Synchronous Verification}

\paragraph{AGREE.}
AGREE~\cite{agree-techreport,agree-observers} is close to Kairos on the logical
structure of specifications. It emphasizes assumptions and guarantees attached
to components, and it supports observer-style encodings of temporal patterns.
This places it closer to Kairos than Aorai with respect to contract structure.
However, AGREE works at an architectural and compositional level, not at the
level of an explicit semantic product between a program automaton and two
safety automata. Kairos differs by making the local product step itself the
primary justification point for local clauses. AGREE is therefore close to
Kairos on the shape of specifications, but not on the proof-theoretic role
played by explicit automata/program synchronization.

\paragraph{Kind 2 and CoCoSpec.}
Kind~2~\cite{kind2-realizability} and CoCoSpec~\cite{cocospec} are also strong
reference points for contract-based reasoning on synchronous systems. Kind~2
offers realizability checking and proof support for Lustre-style programs and
contracts, while CoCoSpec provides a mode-aware contract language for reactive
systems. These systems are close to Kairos in their emphasis on reactive
contracts and explicit assumptions/guarantees. The main difference is again the
central semantic object. Kairos builds its local proof story around the product
between a program, an assumption automaton, and a guarantee automaton. Kind~2
and CoCoSpec are closer to invariant generation, model checking, and contract
reasoning directly over synchronous transition systems than to an explicit local
matching relation between concrete ticks and abstract product steps. This is an
important distinction: Kairos does not seek only an invariant-based proof over
the source transition system, but a semantic bridge from stream-level safety to
local product-step reasoning.

\paragraph{Lustre and theorem proving.}
Older work on Lustre and theorem proving~\cite{lustre-pvs} is also relevant as
historical background. It already combines a reactive programming language with
machine-checked reasoning and therefore shares with Kairos the ambition of
giving a mathematical semantics that supports actual proofs. Kairos differs by
organizing the reduction around safety automata and locally generated
clauses and triples, rather than around direct theorem proving over the source
transition system alone. In that sense, Kairos is closer to a proof-oriented
compilation pipeline than to a direct embedding of the source semantics into a
proof assistant.

\subsection{Monitoring and Observer Compilation}

\paragraph{Copilot.}
Copilot~\cite{copilot-github,copilot3} is an important point of comparison on
the compilation side. It compiles temporal and stream-oriented specifications
into executable monitors, typically generating hard real-time C code. This is
close to Kairos in the sense that both approaches turn temporal specifications
into finite operational artifacts. The main difference is that Copilot is first
and foremost a runtime verification framework, whereas Kairos uses its compiled
artifacts to generate proof objects for deductive verification.
Copilot is particularly relevant as related work on monitor compilation, but it
is further from Kairos than Aorai, AGREE, or Kind~2 on the proof-theoretic
side.

\paragraph{Summary of the comparison.}
The literature already contains, separately, deductive encodings of temporal
properties over programs, synchronous contract reasoning, and compilation of
temporal monitors. What is rarer is a mathematically explicit account of how a
global conditional safety theorem over reactive streams is reduced to local
proof objects when the temporal layer is finite-state but the program itself is
not. This is the niche filled by the present Kairos formulation.

\subsection{Position of Kairos}

Relative to these systems, Kairos combines four ingredients that are rarely
present together in a single framework:
\begin{itemize}[leftmargin=1.4em]
\item a reactive program semantics on streams;
\item an explicit input assumption automaton and an explicit guarantee
  automaton;
\item a semantic product \(program \times A \times G\) used as the source of
  local clauses;
\item a proof-oriented reduction from global bad executions to falsified local
  safety clauses and then to contradictory valid triples.
\end{itemize}
This makes Kairos simultaneously close to annotation-based temporal program
verification, to assume/guarantee reasoning for synchronous systems, and to the
observer-compilation tradition. What distinguishes it is not simply the use of
automata, nor simply the use of deductive verification, but the fact that the
explicit product is the semantic source of generated clauses and triples, while
the program itself is allowed to live over unbounded domains. Put differently,
Kairos occupies a specific point in the design space: it inherits the temporal
discipline of finite safety automata, the proof ambition of deductive program
verification, and the contract structure of synchronous assume/guarantee
reasoning, but it organizes them around an explicit product semantics rather
than around a purely operational monitor or a purely invariant-oriented
transition-system proof.

\section{Discussion}

\subsection{Why One Must Constantly Return to the Product}

A major source of confusion in this kind of document is to talk about
clauses as if they were floating formulas. In reality, every Kairos
clause must always be read as the projection of some information carried by
the product:
\begin{itemize}[leftmargin=1.4em]
\item a reachable product state;
\item a safe product step;
\item a dangerous product step;
\item or a user annotation reattached to that state or step.
\end{itemize}
This discipline explains the editorial choice of the present text: it returns
systematically to the program and to the product before discussing the backend.

\subsection{Limits of the Present Note}

This document does not claim to replace:
\begin{itemize}[leftmargin=1.4em]
\item the detailed Rocq proof;
\item the exhaustive description of the Why3/OBC backend;
\item or a complete complexity study.
\end{itemize}
Its aim is more focused: to provide a mathematical presentation that is complete
and pedagogical enough to serve as a stable reference between the proof, the
code, and the documentation.

\section{Conclusion}

The right central object for Kairos is the explicit product
\[
\text{program} \times A \times G.
\]
It makes the following simultaneously visible:
\begin{itemize}[leftmargin=1.4em]
\item the reactive semantics of the program on streams;
\item the reading of assumptions and guarantees by safety automata;
\item the notion of dangerous step;
\item the generation of local clauses and triples;
\item and the bridge both to the Rocq proof and to the external backend.
\end{itemize}

The main methodological lesson is therefore precise. In a high-quality account
of Kairos, clauses should never be introduced as a mere list of local formulas.
They must always appear as projections of the explicit product, and the actual
proof objects exported to deductive verification must be relational Hoare
triples assembled from those projections. This is what keeps the meaning of the
final theorem stable while moving from stream-level safety to local proof
objects.

Just as importantly, the paper clarifies the scope of what has and has not been
proved. The core theorem concerns reactive semantics, generated clauses,
generated relational triples, and their validity. Backend grouping, Why3
encoding, and solver integration remain refinement layers. This separation is
not a weakness of the presentation; it is part of the scientific contribution,
because it identifies exactly which part of the architecture carries the
semantic content of the correctness argument.

The remaining research directions are therefore well delimited: strengthen the
mathematical account of backend refinement, characterize more precisely the
well-formed program fragment accepted by the implementation, and study how far
the present clause-and-triple discipline can be pushed beyond the current
deterministic safety setting.

\begin{thebibliography}{10}

\bibitem{stouls-groslambert-2009}
Julien Groslambert and Nicolas Stouls.
\newblock V\'erification de propri\'et\'es LTL sur des programmes C par
  g\'en\'eration d'annotations.
\newblock In \emph{Approches Formelles dans l'Assistance au D\'eveloppement de
  Logiciels (AFADL)}, 2009.

\bibitem{aorai-manual}
Nicolas Stouls.
\newblock \emph{Aora\"\i{} Plug-in Tutorial}.
\newblock Frama-C manual, available from the Frama-C documentation site.

\bibitem{cafe-page}
Frama-C Team.
\newblock \emph{CaFE (CaRet Frama-C Extension)}.
\newblock Frama-C plugin documentation page.

\bibitem{agree-techreport}
John D. Backes, Andrew Gacek, Michael W. Whalen, Lee Pike, and others.
\newblock \emph{AGREE: Assume-Guarantee Reasoning Environment}.
\newblock Technical report, Loonwerks, 2021.

\bibitem{agree-observers}
John D. Backes, Michael W. Whalen, Andrew Gacek, and John Komp.
\newblock On implementing real-time specification patterns using observers.
\newblock In \emph{NASA Formal Methods}, 2016.

\bibitem{kind2-realizability}
Temesghen Kahsai, Andrew Gacek, Yeting Ge, and others.
\newblock Realizability checking of contracts with Kind 2.
\newblock In \emph{NASA Formal Methods}, 2022.

\bibitem{cocospec}
Curtis et al.
\newblock CoCoSpec: a mode-aware contract language for reactive systems.
\newblock In \emph{Formal Methods in Computer-Aided Design}, 2016.

\bibitem{lustre-pvs}
Saddek Bensalem, Pascale Caspi, David Chemouil, and others.
\newblock A methodology for proving control systems with Lustre and PVS.
\newblock In \emph{Proceedings of the 7th International Workshop on Higher
  Order Logic Theorem Proving and Its Applications}, 1994.

\bibitem{copilot3}
Lee Pike, Alwyn Goodloe, Jeffrey Maddalon, and others.
\newblock Copilot 3: monitoring autonomous systems.
\newblock NASA technical memorandum, 2020.

\bibitem{copilot-github}
Copilot Language Project.
\newblock \emph{Copilot: a stream-based runtime verification framework}.
\newblock Project repository and documentation.

\bibitem{kupferman-vardi-safety}
Orna Kupferman and Moshe Y. Vardi.
\newblock Model checking of safety properties.
\newblock In \emph{Computer Aided Verification}, 1999.

\bibitem{gastin-oddoux}
Paul Gastin and Denis Oddoux.
\newblock Fast LTL to B{\"u}chi automata translation.
\newblock In \emph{Computer Aided Verification}, 2001.

\bibitem{why3}
Fran{\c c}ois Bobot, Jean-Christophe Filli{\^a}tre, Claude March{\'e}, and
Andrei Paskevich.
\newblock Why3: shepherd your herd of provers.
\newblock In \emph{Boogie}, 2011.

\end{thebibliography}

\end{document}
